---
title: "Evaluating Synthetic Population Quality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Evaluating Synthetic Population Quality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
```{r results='hide', message=FALSE, warning=F}
library(replica)
library(data.table)
```

**Note - the replica library is currently very experimental. It is still undergoing active development and testing. Function, class and method names, arguments and syntax may change without the use of deprecation conventions. This library should currently be used only for exploratory purposes.**


## Motivation

Generating a synthetic population is only part of the population-synthesis workflow.

It is equally important to assess whether the resulting population adequately reproduces the demographic distributions used during construction.

The `replica` package provides tools for:

- comparing observed and expected distributions;

- calculating goodness-of-fit statistics;

- identifying poorly fitted demographic groups; and

- visualising synthetic population quality.

This vignette demonstrates how to:

- inspect validation results;

- compare observed and expected distributions;

- interpret validation statistics; and

- visualise population quality.

## Workflow overview
This vignette focuses on the final stage of the replica workflow.

Aggregate Counts
       ↓
make_agents()
       ↓
Synthetic Agents
       ↓
ConditionalAttributeAdder
       ↓
Enriched Population
       ↓
HouseholdGrouper
       ↓
Synthetic Households
       ↓
Validation


## Validation checklist
Throughout this vignette we will evaluate fit using three complementary diagnostics:

- observed versus expected distributions;

- percentage-point differences;

- goodness-of-fit statistics.

Together these provide both visual and statistical evidence about synthetic population quality.

##  Example data generation process
We begin by creating a simple synthetic population.


```{r}
population <- data.table(

  age_group = c(
    rep("18-64", 20),
    rep("65+", 20)
  ),

  gender = c(
    rep(c("Male", "Female"), 10),
    rep(c("Male", "Female"), 10)
  )

)
```

The synthetic population will be assigned education values using the following contingency table.

```{r}
contingency <- data.frame(

  age_group = c(
    "18-64", "18-64",
    "18-64", "18-64",
    "65+",   "65+",
    "65+",   "65+"
  ),

  gender = c(
    "Male",   "Male",
    "Female", "Female",
    "Male",   "Male",
    "Female", "Female"
  ),

  education = c(
    "Degree", "School",
    "Degree", "School",
    "Degree", "School",
    "Degree", "School"
  ),

  count = c(
    60, 40,
    55, 45,
    30, 70,
    25, 75
  )

)
```

A synthetic population with an education variable can be created using `ConditionalAttributeAdder`.

```{r}
adder <- ConditionalAttributeAdder(

  synth_pop = population,

  contingency = contingency,

  target_attribute = "education",

  group_by = c(
    "age_group",
    "gender"
  )

)

adder <- run(
  adder
)

```

The resulting synthetic population can be summarised using:

```{r}
observed <- synthetic_population_to_contingency(

    adder@synth_pop,
    c(
      "age_group",
      "gender",
      "education"
    )
  )

observed
```

The observed contingency table describes the distribution that actually exists in the synthetic population.


## Visualising validation results

As described in another vignette, validation results are stored in `adder@validation_results`. In this vignette we will outline how these results can be explored through visualisation.

### Visualising observed synthetic and expected distributions

Observed synthetic and expected percentages can be compared graphically.

```{r}
plot_validation_distributions(
  adder@validation_results
)
```

This is usually the first plot to inspect.

Large differences between the observed and expected bars indicate that the attribute assignment process has not fully
reproduced the target distribution.

In this example the bars are generally similar, suggesting that the synthetic population reproduces the reference
distribution reasonably well. 

### Visualising percentage-point differences

Percentage-point differences are often easier to interpret than raw counts.

```{r}
plot_validation_differences(
  adder@validation_results
)
```

The percentage-point difference plot highlights mismatches more clearly than the distribution plot.

Values above zero indicate over-representation.

Values below zero indicate under-representation.

As a rule of thumb:

- differences smaller than 5 percentage points are often
  negligible;
  
- differences between 5 and 10 percentage points may warrant
  investigation;
  
- larger differences suggest a potentially important
  discrepancy.

### Visualising overall fit

For larger contingency tables, a heatmap can provide a useful overview of fit quality.

```{r}
plot_validation_heatmap(
  adder@validation_results
)
```

The heatmap becomes increasingly useful as the number of conditioning variables grows.

For small contingency tables (like those used in this example) the previous figures are often easier to interpret.

For high-dimensional validation exercises the heatmap provides a compact summary of fit quality across many groups
simultaneously.

## Bad fits
To illustrate a poor fit, suppose that all education values in the synthetic population were incorrectly assigned as
`Degree`.

The resulting validation plots would show substantial differences between observed and expected distributions, large
percentage-point differences, and a poor goodness-of-fit statistic.

## Key takeaways

In this vignette we:

- summarised a synthetic population using contingency tables;

- compared observed and expected distributions;

- examined percentage-point differences;

- visualised validation results; and

 - used stored validation diagnostics.

Together, these tools provide a practical framework for assessing the quality of synthetic populations generated using replica.