---
title: "replica"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{replica}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(replica)
```

## Motivation
`replica` is an R package for generating synthetic populations of individual agents and households from aggregated data.

The package has been developed to support health-economic microsimulation modelling and other applications requiring realistic synthetic populations.

`replica` provides tools for:

- creating synthetic agents from aggregate count data;

- assigning demographic and behavioural attributes using contingency tables;

- generating synthetic households; and

- evaluating synthetic population quality.

`replica` implements an R version of the synthetic population generation methodology described by de Mooij et al. (2024) and additionally provides functionality for validation, visualisation and workflow integration within the R ecosystem.

## Status
This development version of `replica` has been made available as part of the process of testing and documenting the library. 

**Note - the `replica` library is currently very experimental. It is still undergoing active development and testing. Function, class and method names, arguments and syntax may change without the use of deprecation conventions. This library should currently be used only for exploratory purposes.**

## The Population Generation Workflow

The workflow supported by `replica` can be summarised as:

Aggregate Counts
       ↓
`make_agents()`
       ↓
Synthetic Agents
       ↓
`ReplicaAdder`
       ↓
Enriched Population
       ↓
`ReplicaStructure`
       +
`ReplicaGrouper`
       ↓
Synthetic Households
       ↓
Validation


Each stage is described in a dedicated vignette.

## Core Concepts

### Synthetic Agents
Synthetic agents are individual records representing people within a synthetic population.

The `make_agents()` function expands aggregate demographic counts into one row per synthetic individual.

Example:

```{r}
age_group <- data.frame(
  age_group = c(
    "0-17",
    "18-64",
    "65+"
  ),
  count = c(
    200,
    600,
    200
  )
)

agents <- make_agents(
  age_group
)
```


### Attribute Assignment

Synthetic populations often require additional characteristics beyond those available in the source data.

`ReplicaAdder` assigns new attributes while preserving known demographic relationships.

Examples include:

- education;

- employment status;

- occupation;

- income category; and

- health status.


### Household generation
Many modelling applications require individuals to be organised into households.

Household generation is controlled using:

- `ReplicaStructure`

- `ReplicaGrouper`

Household structures are generated while respecting demographic and geographic constraints.

### Validation

Synthetic populations should be evaluated to ensure they reproduce the source distributions used during construction.

`replica` provides tools for:

- contingency-table reconstruction;

- goodness-of-fit statistics;

- validation diagnostics; and

- validation visualisations.

## Vignettes

The package documentation follows the complete synthetic population generation workflow.

[Vignette 1: Generating Synthetic Populations from Aggregated Data](V_01.html): 

- introduces `make_agents()`; and

- demonstrates how aggregate count data can be converted into individual synthetic agents.

[Vignette 2: Assigning Attributes Using Contingency Tables](V_02.html):

- introduces `ReplicaAdder`; and

- demonstrates how to enrich synthetic agents using contingency tables and demographic distributions.

[Vignette 3: Generating Synthetic Households](V_03.html):

- introduces `ReplicaStructure` and `ReplicaGrouper`; and

- demonstrates how enriched agents can be organised into realistic household structures.

[Vignette 4: Evaluating Synthetic Population Quality](V_04.html):

- introduces `validate_synthetic_population_fit()`, `plot_validation_distributions()`, `plot_validation_differences()` and `plot_validation_heatmap()`;

- demonstrates how synthetic populations can be assessed and validated.

## Example Workflow

The following example illustrates how the major components of `replica` fit together.

### Create Synthetic Agents

```{r}
age_gender <- data.frame(
  age_group = c("18-64","18-64","65+","65+"),
  gender = c("Male","Female","Male","Female"),
  count = c(10, 10, 10, 10))
agents <- make_agents(age_gender)
```

## Assign Additional Attributes

```{r}
education_table <- data.frame(
  age_group = c("18-64", "18-64", "18-64", "18-64",
                "65+",   "65+",  "65+",   "65+"),
  gender = c("Male", "Male", "Female", "Female",
             "Male", "Male", "Female", "Female"),
  education = c("Degree", "School","Degree", "School",
                "Degree", "School","Degree", "School"),
  count = c(60, 40, 55, 45, 30, 70, 25, 75)

)
adder <- ReplicaAdder(
  population = agents,
  contingency_table = education_table,
  target_attribute = "education",
  group_by = c("age_group","gender")

)
adder <- enhance(adder)
population <- adder@population
```

### Generate Households

```{r}
population[,neighb_code := "N1"]

population[,household_position := "Parent"]

population[age_group == "18-64",
           age := sample(18:64, .N, replace = TRUE)]

population[age_group == "65+", 
           age := sample(65:95,.N, replace = TRUE)]

hh <- ReplicaStructure("CoupleHousehold")

hh <- renew(hh,
            what = "positions",
            household_position = "Parent",
            position_identifier = "adult",
            amount = 2,
            backup_position_identifiers = character())

hh@couple_gender_distribution <- c("Female|Male" = 1)

hh@couple_age_distribution <- c("-5-5" = 1)

hg <- ReplicaGrouper(population = population,
                       group_by = "neighb_code")

hg <- renew(hg, hh)

households <- manufacture(hg)
```

### Inspect Results

```{r}
head(households$synthetic_population)
```

```{r}
head(households$synthetic_households)
```

### Validate Attribute Assignment

```{r}
adder@validation_results[c("z_square","p_value", 
                           "warning_required")]
```

```{r}
plot_validation_differences(adder@validation_results)
```

This workflow demonstrates the complete progression from aggregate demographic data to synthetic agents, enriched populations, synthetic households and validation outputs.

## References

de Mooij J, Sonnenschein T, Pellegrino M, Dastani M, Ettema D, Logan B and Verstegen JA (2024).

*GenSynthPop: generating a spatially explicit synthetic population of individuals and households from aggregated data.*

Autonomous Agents and Multi-Agent Systems. https://link.springer.com/article/10.1007/s10458-024-09680-7
