---
title: "Generating Synthetic Households"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Synthetic Households}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
```{r results='hide', message=FALSE, warning=F}
library(replica)
library(data.table)
```

**Note - the replica library is currently very experimental. It is still undergoing active development and testing. It should be used only for exploratory purposes.**

## Motivation

Many synthetic population generation workflows produce individual agents but do not generate the household relationships needed by simulation models.

For example, health-economic microsimulation models often require agents to be grouped into households before further analysis can be undertaken.

The `replica` package provides tools for transforming an agent-level synthetic population into a collection of synthetic households while preserving demographic and
geographic constraints. To do so `replica`  implements an R port of a third-party python algorithm [developed and described by Jan de Mooij, Tabea Sonnenschein, Marco Pellegrino, Mehdi Dastani, Dick Ettema, Brian Logan & Judith A. Verstegen](https://link.springer.com/article/10.1007/s10458-024-09680-7). 

Household generation in `replica` is based on two core classes:

- `HouseholdType`

- `HouseholdGrouper`

A `HouseholdType` describes a household structure that should be generated, while a `HouseholdGrouper` applies one or more household definitions to a synthetic population.

This vignette demonstrates the complete workflow for creating synthetic households.

## Supply a synthetic population of individual agents
This vignette assumes that we have already created a synthetic population of individual agents. The following code generates an illustrative example population of individual agents. For simplicity, this example contains four adults living in the same neighbourhood.

```{r}
pop <- data.table(

  agent_id = c(
    "A001",
    "A002",
    "A003",
    "A004"
  ),

  neighb_code = c(
    "N1",
    "N1",
    "N1",
    "N1"
  ),

  age = c(
    40,
    35,
    50,
    45
  ),

  gender = c(
    "Male",
    "Female",
    "Male",
    "Female"
  ),

  household_position = c(
    "Parent",
    "Parent",
    "Parent",
    "Parent"
  )

)
```

The `household_position` column is particularly important. 

Rather than storing household membership directly, `replica` uses household-position categories to indicate the types of roles each agent may occupy within a household.

Examples include:

- `Parent`
- `Child`
- `Grandparent`
- `Student`

These positions are subsequently referenced when household types are defined. In this example all agents are potential parents.

## Create a household type
A household type specifies a household structure that we plan to generate. Here we create a simple couple household.

```{r}
hh <- HouseholdType(
  "CoupleHousehold"
)

```

At this point we have only defined the household name. We now need to specify the household members that form this household type.


## Define household membership

The addMembers() method defines the roles that must be filled when constructing a household. In this example, a couple household requires two adults.

```{r}
hh <- addMembers(
  hh,
  household_position = "Parent",
  position_identifier = "adult",
  amount = 2,
  backup_position_identifiers = character()
)
```

This tells replica that:

 - agents with household position "Parent" are eligible;
 
 - these agents represent "adult" household members; and
  
 - exactly two adults are required per household.
 
 
 Notice that two separate concepts are used:

- `household_position`

- `position_identifier`

The `household_position` corresponds to values that appear in the synthetic population.

The `position_identifier` defines broader household roles used internally by the household-building algorithm.

For example, multiple population-specific household positions could all be assigned the identifier `"adult"`.

This separation makes it easier to adapt household generation rules to different source datasets.

The household definition is now complete.


## Define houshold matching rules
Household construction often requires demographic matching rules. For example:

- for a given household type, should couples be opposite-sex or same-sex?

 - what age differences should be used to match partners?

These rules can be specified using distributions.

For simplicity, this illustrative example will generate Female - Male pairs in which the age gap between partners is within 5 years.

```{r}
hh@couple_gender_distribution <- c("Female|Male" = 1)

hh@couple_age_distribution <- c("-5-5" = 1)
```

The above age distribution indicates that permissible partner age gaps range from five years younger to five years older.

## Create a household grouper

A HouseholdGrouper coordinates household generation across an entire synthetic population.

```{r}
hg <- HouseholdGrouper(df_synth_pop = pop,
                       group_by = "neighb_code")
```

The group_by argument determines the geographic or administrative boundaries within which household generation will occur.

In this example households are generated independently within each neighbourhood.

## Register household types

A HouseholdGrouper may contain multiple household types.

For example, a realistic population might contain:

 - single-adult households;

 - couple households; and

 - family households.

Household types are registered using addHouseholdType().

```{r}
hg <- addHouseholdType(hg,hh)
```

The grouper now knows how to generate couple households.

## Run household generation

Once household types have been registered, the complete household-generation workflow can be executed using run().

```{r}
result <- run(hg)
```

The `run()` method executes the complete household-generation workflow.

For each grouping region (`neighb_code` in this example), `replica`:

1. Identifies eligible household members.
2. Creates household structures using registered `HouseholdType` objects.
3. Matches adults according to the configured gender and age distributions.
4. Creates synthetic households.
5. Assigns household identifiers.
6. Produces household-level summary tables.

The result is returned as a list containing both agent-level and household-level outputs.
 
## Inspect output
The result now contains both agent-level and household-level outputs. 

```{r}
synthetic_population <- result$synthetic_population
```

```{r}
synthetic_households <- result$synthetic_households
```

The household-generation workflow produces two complementary outputs.

`synthetic_population` contains one row per agent and records the household assignment of each synthetic individual.

`synthetic_households` contains one row per household and summarises the resulting household structures.

Together these outputs allow analyses to be performed at both the individual and household level.

### Inspect household assignments

The agent-level output now contains household identifiers.

```{r}
synthetic_population[,c("agent_id", "household_id")]
```

Agents sharing the same household identifier belong to the same synthetic household.

### Inspec household-level summaries
The household summary table provides one row per household.

```{r}
synthetic_households
```

This table is often useful when producing reports or validating generated household structures.

## Validate household integrity

The household-generation workflow automatically performs validation checks.

Additional validation can be performed manually.

```{r}
any(
  is.na(
    synthetic_population$household_id
  )
)
```

A successful integrity check indicates that:

 - no agents have been assigned more than once;
 
 - no eligible agents have been omitted;
 
  - household membership records are internally consistent.
  
## Create family households

The previous example generated households containing adults only.

In practice, many synthetic populations will require more complex household structures containing both adults and children.

The same workflow can be extended by registering additional member roles.

```{r}
hh <- HouseholdType("FamilyHousehold")
```

```{r}
hh <- addMembers(
  hh,
  household_position = "Parent",
  position_identifier = "adult",
  amount = 2,
  backup_position_identifiers = character()
)

hh <- addMembers(
  hh,
  household_position = "Child",
  position_identifier = "child",
  amount = 2,
  backup_position_identifiers = character()
)
```


When child household members are present, `replica` automatically:

- forms sibling groups;

- creates adult partner groups;

- matches adults and children; and

- creates complete family households.

No additional user intervention is required.


The resulting household structures are stored using separate adult, child and all components.


## How the workflow fits together

The overall workflow can be summarised as:

HouseholdType
      ↓
addMembers
      ↓
HouseholdGrouper
      ↓
addHouseholdType
      ↓
run
      ↓
Synthetic Population
      +
Synthetic Households


## Key takeaways

In this vignette we:

- created a synthetic population;

- defined a household structure using `HouseholdType`;

- specified adult household members;

- configured demographic matching rules;

- registered a household type using `HouseholdGrouper`;

- generated synthetic households using `run()`; and

- explored agent-level and household-level outputs.

The same workflow can be extended to support multiple household types, family structures, and more complex demographic matching requirements.
