---
title: "Generating Synthetic Populations from Aggregated Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Synthetic Populations from Aggregated Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
```{r results='hide', message=FALSE, warning=F}
library(replica)
```

**Note - the replica library is currently very experimental. It is still undergoing active development and testing. Function, class and method names, arguments and syntax may change without the use of deprecation conventions. This library should currently be used only for exploratory purposes.**

## Motivation
Synthetic population generation often begins with aggregate population statistics rather than individual-level records.

Examples include:

- population counts by neighbourhood;

- age-group marginals;

- demographic contingency tables;

- census tabulations.

The first step in many population-synthesis workflows is to convert these aggregate counts into a synthetic population of individual agents.

The `make_agents()` function performs this task by expanding aggregate counts into one row per synthetic individual. This function aims to implement in R the initial step of an algorithm described by de Mooij et al. (2024). 

This vignette demonstrates how to create synthetic agents from marginal distributions and contingency tables and shows how these agents can subsequently be enriched and grouped using the remainder of the `replica` workflow. 

## Introduction
The workflow supported by `replica` can be summarised as:

Aggregate Counts
       
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; ↓
       
`make_agents()`
       
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; ↓
       
Synthetic Agents
       
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; ↓
       
`ReplicaAdder`
       
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; ↓

`ReplicaStructure`

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; +

`ReplicaGrouper`
      
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; ↓
       
Validation

This vignette focuses on the first stage: creating a synthetic population of individual agents.



## Example 1: Creating agents from a spatial marginal
Suppose we know the number of residents in each neighbourhood.


```{r}
neighbourhoods <- data.frame(neighb_code = c( "N1", "N2", "N3" ), 
                            count = c( 100, 150, 75 ) ) 
neighbourhoods
```

The table contains aggregate counts rather than individual agents. We can convert this table to a table of individual agents:

```{r}
agents <- make_agents(
  neighbourhoods
)

head(agents)
```

Each row now represents a synthetic individual. 

We can confirm the number of generated agents:

```{r}
nrow(agents)
```
The generated population can also be aggregated back into a contingency table.


```{r}
transform_to_contingency(

  agents,

  "neighb_code"

)
```

The reconstructed counts should match the source table used to generate the agents.

## Example 2: Creating agents from an age marginal
The same function can be used with any marginal distribution.

```{r}
age_margin <- data.frame(

  age_group = c(
    "0-17",
    "18-64",
    "65+"
  ),

  count = c(
    200,
    600,
    200
  )

)

age_margin
```

```{r}
agents <- make_agents(
  age_margin
)

head(agents)
```

Each synthetic agent now contains an age-group assignment.

Summarise the resulting population:

```{r}
transform_to_contingency(

  agents,

  "age_group"

)
```

## Example 3: Creating agents from a contingency table

More detailed synthetic populations can be generated from contingency tables.

```{r}
age_gender <- data.frame(

  age_group = c(
    "0-17",
    "0-17",
    "18-64",
    "18-64",
    "65+",
    "65+"
  ),

  gender = c(
    "Male",
    "Female",
    "Male",
    "Female",
    "Male",
    "Female"
  ),

  count = c(
    100,
    90,
    310,
    290,
    110,
    100
  )

)

age_gender
```

Generate agents:

```{r}
agents <- make_agents(
  age_gender
)
```

Inspect the resulting population:

```{r}
head(agents)
```

We can compare the generated population with the original contingency table.

```{r}
transform_to_contingency(agents, 
                                    columns = c("age_group", "gender"))
```

The reconstructed contingency table should reproduce the original counts exactly.

This property makes make_agents() a useful foundation for more sophisticated synthetic-population workflows.

## Add agent attributes
The agents created by make_agents() typically contain only a subset of the characteristics required by a simulation model.

Additional attributes can be assigned using `ReplicaAdder`. The use of this class is demonstrated in the "Assigning Attributes Using Contingency Tables" vignette.

## Assign agents to households
After additional attributes have been assigned, agents can be grouped into households.

To do this, use the `ReplicaStructure` and `ReplicaGrouper` classes. These are described in another vignette.


## Key takeaways

In this vignette we:

- introduced synthetic-agent generation;

- created agents from spatial marginals;

- created agents from demographic marginals;

- created agents from contingency tables;

- verified that aggregate distributions were preserved;

- connected agent generation to the broader `replica` workflow.


## References

de Mooij J, Sonnenschein T, Pellegrino M, Dastani M, Ettema D, Logan B and Verstegen JA (2024).

*GenSynthPop: generating a spatially explicit synthetic population of individuals and households from aggregated data.*

Autonomous Agents and Multi-Agent Systems. https://link.springer.com/article/10.1007/s10458-024-09680-7

