---
title: "Generating Synthetic Households with replica"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Synthetic Households with replica}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
```{r results='hide', message=FALSE, warning=F}
library(replica)
library(data.table)
```

**Note - the replica library is currently very experimental. It is still undergoing active development and testing. It should be used only for exploratory purposes.**

## Generating synthetic households with replica

The `replica` package provides tools for constructing synthetic households for use in mental health economic microsimulation models.

Currently, only one method for generating synthetic populations is supported - which is an R port of a third-party python algorithm [developed and described by Jan de Mooij, Tabea Sonnenschein, Marco Pellegrino, Mehdi Dastani, Dick Ettema, Brian Logan & Judith A. Verstegen](https://link.springer.com/article/10.1007/s10458-024-09680-7). This vignette describes the household generation component of this algorithm as implemented by replica. 

Household generation is performed using two core classes:

- `HouseholdType`
- `HouseholdGrouper`

A `HouseholdType` defines a household structure (for example, a couple household or family household), while a `HouseholdGrouper` applies one or more household types to a synthetic population and generates household assignments.

This vignette demonstrates the complete household-generation workflow.

## Supplying a synthetic population of individuals
This vignette assumes that we have already created a synthetic population of individual agents. The following code generates an illustrative example population of individual agents. For simplicity, this example contains four adults living in the same neighbourhood.

```{r}
pop <- data.table(

  agent_id = c(
    "A001",
    "A002",
    "A003",
    "A004"
  ),

  neighb_code = c(
    "N1",
    "N1",
    "N1",
    "N1"
  ),

  age = c(
    40,
    35,
    50,
    45
  ),

  gender = c(
    "Male",
    "Female",
    "Male",
    "Female"
  ),

  household_position = c(
    "Parent",
    "Parent",
    "Parent",
    "Parent"
  )

)
```

Each row represents one synthetic agent.

The household_position column describes the role that each agent may occupy within a household. In this example all agents are potential parents.

## Creating a household type
A household type specifies a household structure that we plan to generate. Here we create a simple couple household.

```{r}
hh <- HouseholdType(
  "CoupleHousehold"
)

```

At this point we have only defined the household name. We now need to specify the household members that form this household type.


## Defining household membership

The addMembers() method defines the roles that must be filled when constructing a household. In this example, a couple household requires two adults.

```{r}
hh <- addMembers(
  hh,
  household_position = "Parent",
  position_identifier = "adult",
  amount = 2,
  backup_position_identifiers = character()
)
```

This tells replica that:

 - agents with household position "Parent" are eligible;
 
 - these agents represent "adult" household members; and
  
 - exactly two adults are required per household.

The household definition is now complete.


## Defining houshold matching rules
Household construction often requires demographic matching rules. For example:

- for a given household type, should couples be opposite-sex or same-sex?

 - what age differences should be used to match partners?

These rules can be specified using distributions.

For simplicity, this illustrative example will generate Female - Male pairs in which the age gap between partners is within 5 years.

```{r}
hh@couple_gender_distribution <- c("Female|Male" = 1)

hh@couple_age_distribution <- c("-5-5" = 1)
```

The above age distribution indicates that permissible partner age gaps range from five years younger to five years older.

## Creating a household grouper

A HouseholdGrouper coordinates household generation across an entire synthetic population.

```{r}
hg <- HouseholdGrouper(df_synth_pop = pop,
                       group_by = "neighb_code")
```

The group_by argument determines the geographic or administrative boundaries within which household generation will occur.

In this example households are generated independently within each neighbourhood.

## Registering household types

A HouseholdGrouper may contain multiple household types.

For example, a realistic population might contain:

 - single-adult households;

 - couple households; and

 - family households.

Household types are registered using addHouseholdType().

```{r}
hg <- addHouseholdType(hg,hh)
```

The grouper now knows how to generate couple households.

## Running household generation

Once household types have been registered, the complete household-generation workflow can be executed using run().

```{r}
result <- run(hg)
```

Internally this workflow:

 - identifies eligible agents;

 - creates household member groups;

 - matches partners;

 - assigns household identifiers;

 - validates household assignments; and

 - creates household-level summary tables.
 
## Inspecting output
The result now contains both agent-level and household-level outputs. 

### Synthetic population

The updated synthetic population is available via:

```{r}
synthetic_population <- result$synthetic_population
```

### Synthetic households

Household summaries are available via:

```{r}
synthetic_households <- result$synthetic_households
```

### Inspecting household assignments

The agent-level output now contains household identifiers.

```{r}
synthetic_population[,c("agent_id", "household_id")]
```

Agents sharing the same household identifier belong to the same synthetic household.

### Inspecting household-level summaries
The household summary table provides one row per household.

```{r}
synthetic_households
```

This table is often useful when producing reports or validating generated household structures.

## Validating Household Integrity

The household-generation workflow automatically performs validation checks.

Additional validation can be performed manually using checkIntegrity().

```{r}
checkIntegrity(result$object@household_types[[1]])
```

A successful integrity check indicates that:

 - no agents have been assigned more than once;
 
 - no eligible agents have been omitted;
 
  - household membership records are internally consistent.
  
## Family Households

The previous example generated couple households only. More complex household structures can also be created.

For example, family households can include both adults and children.


```{r}
hh <- HouseholdType("FamilyHousehold")
```

```{r}
hh <- addMembers(
  hh,
  household_position = "Parent",
  position_identifier = "adult",
  amount = 2,
  backup_position_identifiers = character()
)

hh <- addMembers(
  hh,
  household_position = "Child",
  position_identifier = "child",
  amount = 2,
  backup_position_identifiers = character()
)
```


When family households are generated, replica automatically:

 - groups children into sibling groups;

 - matches sibling groups to adult groups; and

 - creates household records containing both adults and children.

The resulting household structures are stored using separate adult, child and all components.


## How the Workflow Fits Together

The overall workflow can be summarised as:

HouseholdType
      ↓
addMembers
      ↓
HouseholdGrouper
      ↓
addHouseholdType
      ↓
run
      ↓
Synthetic Population
      +
Synthetic Households


Summary

In this vignette we:

 - created a synthetic population;

 - defined a household structure using HouseholdType;

 - specified household member requirements;

 - defined partner-matching rules;

 - registered a household type with a HouseholdGrouper;

 - generated synthetic households using run();
 
 - examined agent-level and household-level outputs;

 - validated the resulting household assignments.

The same workflow can be extended to more complex populations containing multiple household types, children, and richer demographic matching rules.
