---
title: "Assigning Attributes Using Contingency Tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Assigning Attributes Using Contingency Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
```{r results='hide', message=FALSE, warning=F}
library(replica)
library(data.table)
```

**Note - the replica library is currently very experimental. It is still undergoing active development and testing. It should be used only for exploratory purposes.**

## Motivation

Synthetic populations are often generated with a limited set of demographic variables.

Additional attributes often need to be assigned using aggregate data sources such as:

- census tables;

- survey microdata;

- administrative datasets; and

- official projections.

The `ConditionalAttributeAdder` class provides a framework for assigning attributes to synthetic agents while preserving known conditional distributions. 

This vignette demonstrates how to assign a target attribute using a contingency table. The approach demonstrated in this algorithm is an R port of a third-party python algorithm [developed and described by Jan de Mooij, Tabea Sonnenschein, Marco Pellegrino, Mehdi Dastani, Dick Ettema, Brian Logan & Judith A. Verstegen](https://link.springer.com/article/10.1007/s10458-024-09680-7). 

## Supply a seed synthetic population of individual agents

We begin by creating a small synthetic population that lacks the target attribute.

```{r}
population <- data.table(

  age_group = c(
    rep("18-64", 20),
    rep("65+", 20)
  ),

  gender = c(
    rep(c("Male", "Female"), 10),
    rep(c("Male", "Female"), 10)
  )

)
```


The resulting population contains:

```{r}
population[
  ,
  .N,
  by = c(
    "age_group",
    "gender"
  )
]

```

The population contains age and gender information but no education status.

Our goal is to assign education values while preserving a specified conditional distribution.

## Supply a contingency table
The reference distribution is supplied as a contingency table. In real use, we would source contingency tables from high quality data sources representative of our regions of interest. In Australia this would typically be data provided by the Australian Bureau of Statistics. In this example, we will generate an illustrative contingency table using hypothetical values.

```{r}
contingency <- data.frame(

  age_group = c(
    "18-64", "18-64",
    "18-64", "18-64",
    "65+",   "65+",
    "65+",   "65+"
  ),

  gender = c(
    "Male",   "Male",
    "Female", "Female",
    "Male",   "Male",
    "Female", "Female"
  ),

  education = c(
    "Degree", "School",
    "Degree", "School",
    "Degree", "School",
    "Degree", "School"
  ),

  count = c(
    60, 40,
    55, 45,
    30, 70,
    25, 75
  )

)
```

The contingency table defines the expected education distribution within each age-group and gender combination. The `count` column describes the frequency of each combination in the reference data source.

For example:

18-64 Male
  Degree = 60%
  School = 40%

65+ Female
  Degree = 25%
  School = 75%


## Create a ConditionalAttributeAdder

The assignment workflow is controlled using a `ConditionalAttributeAdder`.

```{r}
adder <- ConditionalAttributeAdder(
  synth_pop = population,
  contingency = contingency,
  target_attribute = "education",
  group_by = c(
    "age_group",
    "gender"
  )
)
```

The object contains:

- the synthetic population;

- the contingency table;

- the target attribute to be assigned;

- the conditioning variables.

## Handle missing groups

Real-world contingency tables often do not contain every combination of characteristics found in a synthetic population. For example, a synthetic population may contain demographic groups that do not appear in the source data.

`replica` supports several strategies for handling missing groups. The selected strategy is stored in the `missing_group_strategy` slot.

```{r}
adder@missing_group_strategy
```

In the above example, the strategy is "borrow" (which is selected by default) that borrows information from nearby groups.

Other options are "overall" which uses the overall distribution of the target attribute and "error" which stops execution when a required group is missing.

## Run attribute assignment

Attribute assignment is performed using run().

```{r}
adder <- run(adder)
```

The workflow:

1. Resolves missing contingency groups.

2. Computes group-specific fractions.

3. Converts fractions into agent allocations.

4. Assigns attribute values.

5. Verifies the resulting distribution.

The updated synthetic population is stored in the `synth_pop` slot


## Inspect results
The synthetic population now contains the assigned `education` attribute.

```{r}
head(adder@synth_pop)
```


We can also examine the resulting contingency table.

```{r}
synthetic_population_to_contingency(
  adder@synth_pop,
  c("age_group", "gender","education"))
```

The assigned education values approximately reproduce the proportions defined in the reference contingency table.


## Add marginal constraints

Additional marginal distributions can be supplied when known population totals are available. Margins provide additional information that can be used when reconciling contingency tables with known distributions.

For example:

```{r}
gender_margin <- data.frame(
  gender = c("Male", "Female"),
  count = c(100, 120))
```

```{r}
adder <- addMargins(adder,
                    margins = list(gender_margin),
                    margins_names = list("gender"))
```

Because the object configuration has changed, any previous validation results are automatically cleared.

Running the workflow again regenerates the validation diagnostics.

```{r}
adder <- run(adder)
```


## Inspect validation results

Validation results are stored directly in the `validation_results` slot.

```{r}
names(adder@validation_results)
```

### Test statistics
We can quickly inspect a selection of these results such as z squared and p-value and whether a warning flag has been raised due to poor match between the expected and observed synthetic results.

```{r}
adder@validation_results[c(
  "z_square",
  "p_value",
  "warning_required"
)]
```

The z-squared statistic provides a measure of divergence between observed synthetic and expected distributions. Smaller values indicate a closer match between the synthetic and reference distributions.

The p-value provides a formal goodness-of-fit test. Large p-values indicate that the observed synthetic population is consistent with the target distribution, while small p-values suggest a poorer fit. If p-values are less than 0.05< the `warning_required` flag will be set to `TRUE`.

Because statistical significance is sensitive to population size, p-values should be interpreted alongside the percentage-point differences reported in validation_results$details. 

### Summary table

A table comparing the count and percentage based distributions in the observed synthetic and expected data contains the following columns:

```{r}
names(adder@validation_results$details)
```

In this table:

- count_x is the observed synthetic count;

- count_y is the expected count from the reference table;

- observed_pct is the observed synthetic percentage;

- expected_pct is the expected percentage; and

- difference_pct is the percentage-point difference (which may provide a more intuitive assessment of fit than examining raw counts).

These diagnostics make it easy to identify groups where the assigned synthetic population differs from the target distribution.

```{r}
adder@validation_results$details[
  ,
  .(
    age_group,
    gender,
    education,
    observed_pct,
    expected_pct,
    difference_pct
  )
]
```


### Validation results and object state
Validation results are stored within the ConditionalAttributeAdder object at `adder@validation_results`.

When configuration settings are modified, previous validation results are automatically cleared.

For example:

```{r}
adder <- addMargins(
  adder,
  margins = list(gender_margin),
  margins_names = list("gender")
)

adder@validation_results
```

After re-running the assignment workflow:

```{r}
adder <- run(adder)
adder@validation_results$z_square
```

the validation results are regenerated automatically.

## Key takeaways

In this vignette we:

- created a synthetic population;

- defined a reference contingency table;

- configured a `ConditionalAttributeAdder`;

- assigned a new attribute using `run()`;

- explored strategies for handling missing groups;

- added marginal distributions; and

- verified the resulting synthetic population.

These tools provide a flexible framework for enriching synthetic populations using aggregate demographic data.

